---
  title: "benchmark report for simulations"
  output:
    html_document:
      fig_height: 20
      fig_width: 20
---

# preliminaries

```{r, eval=FALSE}
n_cpu <- 15
sim_name <- 'gfr_3_3_20_42_2'
sim_name <- 'gcd_3_3_20_1_2'
sim_name <- 'isoform_5_5_15_387_1'
```


```{r}
args <- commandArgs(trailingOnly = TRUE)
if (length(args) != 2) {
  stop('Usage: RScript N_CPU EXPERIMENT_STRING')
}
n_cpu <- args[1]
sim_name <- args[2]
```

The simulation string is:

```{r}
sim_name
```

We use the variable `simulation_mode` to build different axes depending on the simulation.
These axes have been tuned by hand.

```{r}
prefix <- strsplit(sim_name, '_')[[1]][1]
simulation_mode <- switch(prefix,
  'isoform' = 'independent',
  'gcd' = 'common',
  'gfr' = 'reference',
  stop(paste0('unrecognized simulation type. Please check the `sim_name`: ',
    sim_name)
  ))
```

The simulation mode is: `R simulation_mode`.

Some more global variables on where to save the figures:

```{r}
base_dir <- '../results/final_figures'
dir.create(base_dir, showWarnings = F)
default_extension <- '.pdf'
```

A theme that seems to save well:

```{r}
theme_hp <- function() {
  theme_cowplot(25) +
    theme(legend.key.size = unit(2, "lines"))
}
```

## load packages

```{r}
.libPaths(c("~/R_library", .libPaths()))
suppressPackageStartupMessages({
library('cowplot')
library('data.table')
library('dplyr')
library('mamabear')
library('parallel')
})
source('gene_common.R')
options(mc.cores = n_cpu)
```

set the number of columns for each grid

```{r}
n_grid_columns <- 2
```

ignore the following when plotting estimated fdr:

```{r}
ignore_fdr <- c('GLFC', 'LFC')
# method_colors_lfc <- c(method_colors, c(LFC = 'black', GLFC = 'dimgray'))
method_colors_lfc <- c(method_colors, c(LFC = '#4393c3', GLFC = '#fddbc7'))
```

# isoform benchmarks

## each method on their own filter

```{r}
small_sal_benchmarks <- readRDS(paste0('../results/', sim_name,
  '/small_sal_benchmarks.rds'))
small_sal_benchmarks <- small_sal_benchmarks[sort(names(small_sal_benchmarks))]
```

For the paper, let's look at only a subset of the things.
In general, we are recommending the likelihood ratio test and aggregation.
Let's remove everything else:

```{r}
to_remove <- c('sleuth.wt', 'sleuthALR.wt', 'ALDEx2.welsh', 'ALDEx2.wilcoxon')
subset_small_benchmarks <- lapply(small_sal_benchmarks,
  function(bench_list) {
    lapply(bench_list,
      function(bench) {
        filter_benchmark(bench, to_remove, join_mode = 'intersect')
      })
  })
```

```{r}
if (simulation_mode == 'reference') {
  sb <- lapply(subset_small_benchmarks,
    function(bench_list) {
      lapply(bench_list,
        function(bench) {
          current_oracle <- bench$oracle
          cat(paste0('before: ', sum(current_oracle$is_de), '\n'))
          current_oracle <- dplyr::mutate(current_oracle,
            is_de = ifelse(is_de & log_fc < eps & -eps < log_fc, FALSE, is_de))
          cat(paste0('after: ', sum(current_oracle$is_de), '\n'))

          new_de_benchmark(bench$original_data, bench$labels, current_oracle,
            de_colors = bench$color_mapping,
            join_mode = 'intersect')
        })
    })
  subset_small_benchmarks <- sb
}
```

Generate all of the FDR versus number of differentially expressed genes

```{r}
subset_nde_plot <- lapply(subset_small_benchmarks,
  function(bench) {
    suppressMessages(fdr_nde_plot(bench)) +
      theme_hp()
  })

current_limits <- switch(simulation_mode,
    independent = coord_cartesian(xlim = c(-0.01, 2250), ylim = c(-0.01, 0.10),
      expand = FALSE),
    common = coord_cartesian(xlim = c(-0.01, 2000), ylim = c(-0.01, 0.15),
      expand = FALSE),
    reference = coord_cartesian(xlim = c(-0.01, 700), ylim = c(-0.01, 0.15),
      expand = FALSE)
  )
subset_nde_plot <- lapply(subset_nde_plot, function(x) x + current_limits)
plot_grid(plotlist = subset_nde_plot, ncol = n_grid_columns)
```

## estimated fdr versus true fdr

The black line is $x = y$.
If you are below the line, you are conservative.
If you are above the line, you are anti-conservative.

```{r}
current_limits <- switch(simulation_mode,
    independent = coord_cartesian(xlim = c(-0.01, 1), ylim = c(-0.01, 1),
      expand = FALSE),
    common = coord_cartesian(xlim = c(-0.01, 1), ylim = c(-0.01, 1),
      expand = FALSE),
    reference = coord_cartesian(xlim = c(-0.01, 1), ylim = c(-0.01, 1),
      expand = FALSE)
  )

subset_fdr_efdr_plot <- lapply(subset_small_benchmarks,
  function(bench) {
    suppressMessages(fdr_efdr_plot(bench)) +
      theme_hp() +
      theme(legend.position = c(0.80, 0.15)) +
      geom_abline(slope = 1, intercept = 0, color = 'black', linetype = 3)
  })
subset_fdr_efdr_plot <- lapply(subset_fdr_efdr_plot,
  function(x) x + current_limits)
plot_grid(plotlist = subset_fdr_efdr_plot, ncol = n_grid_columns)
```

## box plots of estimated fdr versus true fdr

```{r}
subset_bp <- lapply(subset_small_benchmarks,
  function(bench) {
    suppressMessages(boxplot_prep(bench, c(0.01, 0.05, 0.10)))
  })
```

Looking at the estimated fdr versus the true fdr

```{r}
subset_fdr_efdr_boxplot <- lapply(subset_bp,
  function(bp) {
    fdr_efdr_boxplot(bp)
  })
plot_grid(plotlist = subset_fdr_efdr_boxplot, ncol = n_grid_columns)
```

Looking at the estimated fdr versus the power

```{r}
subset_power_efdr_boxplot <- lapply(subset_bp,
  function(bp) {
    power_efdr_boxplot(bp)
  })
plot_grid(plotlist = subset_power_efdr_boxplot, ncol = n_grid_columns)
```

Looking at the estimated fdr versus the accuracy

```{r}
subset_accuracy_efdr_boxplot <- lapply(subset_bp,
  function(bp) {
    ggplot(bp, aes(method, accuracy)) +
      geom_boxplot(aes(color = method)) +
      facet_wrap(~fdr_level)
  })
plot_grid(plotlist = subset_accuracy_efdr_boxplot, ncol = n_grid_columns)
```

Looking at the estimated fdr versus the specificity

```{r}
subset_specificity_efdr_boxplot <- lapply(subset_bp,
  function(bp) {
    ggplot(bp, aes(method, specificity)) +
      geom_boxplot(aes(color = method)) +
      facet_wrap(~fdr_level)
  })
plot_grid(plotlist = subset_specificity_efdr_boxplot, ncol = n_grid_columns)
```

```{r}
subset_fdr_power <- lapply(subset_small_benchmarks,
  function(bench) {
    fdr_power_plot(bench)
  })
plot_grid(plotlist = subset_fdr_power, ncol = n_grid_columns)
```

## fdr/power curves when taking the intersection of filters

Use `sim_filter = TRUE` on isoforms to remove things that we remove isoforms that we expect to be to locally express to detect.

```{r}
subset_fdr <- lapply(subset_small_benchmarks,
  function(bench) {
    suppressMessages(get_fdr(bench, sim_filter = TRUE)$pvals)
  })
```

```{r}
subset_fdr_power_plot <- lapply(subset_fdr,
  function(bench) {
    suppressMessages(fdr_efdr_power_plot(bench, start = 100, jump = 100,
      isolines = FALSE)) +
      theme_hp()
  })

current_limits <- switch(simulation_mode,
    independent = coord_cartesian(xlim = c(-0.01, 0.25), ylim = c(-0.01, 0.32),
      expand = FALSE),
    common = coord_cartesian(xlim = c(-0.01, 0.25), ylim = c(-0.01, 0.30),
      expand = FALSE),
    reference = coord_cartesian(xlim = c(-0.01, 0.20), ylim = c(-0.01, 0.10),
      expand = FALSE)
  )
subset_fdr_power_plot <- lapply(subset_fdr_power_plot,
  function(x) x + current_limits)
p <- plot_grid(plotlist = subset_fdr_power_plot, ncol = n_grid_columns)
p

filename <- file.path(base_dir, paste0('small.pairwise_zoom_', sim_name,
  default_extension))
save_plot(filename, p, base_aspect_ratio = 1.6, base_height = 15)
```

### The full benchmarks

```{r}
current_limits <- switch(simulation_mode,
    independent = coord_cartesian(xlim = c(-0.01, 2000), ylim = c(-0.01, 0.15),
      expand = FALSE),
    common = coord_cartesian(xlim = c(-0.01, 2000), ylim = c(-0.01, 0.15),
      expand = FALSE),
    reference = coord_cartesian(xlim = c(-0.01, 700), ylim = c(-0.01, 0.10),
      expand = FALSE)
  )

all_nde_plot <- lapply(small_sal_benchmarks,
  function(bench) {
    suppressMessages(fdr_nde_plot(bench)) +
      theme_hp()
  })
all_nde_plot <- lapply(all_nde_plot, function(x) x + current_limits)
plot_grid(plotlist = all_nde_plot, ncol = n_grid_columns)
```

```{r}
all_fdr_efdr_plot <- lapply(small_sal_benchmarks,
  function(bench) {
    suppressMessages(fdr_efdr_plot(bench)) +
      theme_hp() +
      geom_abline(slope = 1, intercept = 0, color = 'black', linetype = 3)
  })
plot_grid(plotlist = all_fdr_efdr_plot, ncol = n_grid_columns)
```

```{r}
all_fdr_power <- lapply(small_sal_benchmarks,
  function(bench) {
    fdr_power_plot(bench, sim_filter = TRUE) +
    theme_hp()
  })

current_limits <- switch(simulation_mode,
    independent = coord_cartesian(xlim = c(-0.01, 0.25), ylim = c(-0.01, 0.10),
      expand = FALSE),
    common = coord_cartesian(xlim = c(-0.01, 0.10), ylim = c(-0.01, 0.10),
      expand = FALSE),
    reference = coord_cartesian(xlim = c(-0.01, 0.10), ylim = c(-0.01, 0.03),
      expand = FALSE)
  )
all_fdr_power <- lapply(all_fdr_power, function(x) x + current_limits)
plot_grid(plotlist = all_fdr_power, ncol = n_grid_columns)
```

--------------------------------------------------------------------------------

# session

```{r}
sessionInfo()
```
